An Improved Method for Coupling Hydrodynamics with Astrophysical Reaction Networks

### 2022-03-26 simplified_sdc_intro/paper.tex
rename plots
fix conserved state
add melissa
add some net details

@@

We solve the Euler equations for compressible, reacting flow.  For ease
of exposition we describe the one-dimensional case;
multidimensional extensions are a straightforward modification to
include the CTU hydrodynamics scheme.  Our conserved variables are
\begin{equation}
  \Uc = \left ( \begin{array}{c}
           \rho \\
           \rho X_k \\
           \rho \alpha_l \\
           \rho u \\
           \rho E \\
           \rho e \end{array}\right )
\end{equation}
where $\rho$ is the mass density, $u$ is velocity, $E$ is specific
total energy, $p$ is the pressure, and we carry nuclear species mass
fractions, $X_k$.  The specific total
energy relates to the specific internal energy, $e$, as $E = e + u^2/2$,
and we also separately evolve $e$, as part of a dual
energy formulation (see \citealt{bryan:1995,wdmergerI}).
 The mass fractions are constrained to sum to 1, $\sum_k X_k = 1$, but
 no such constraint exists on the auxiliary variables. 
Defining the hydrodynamical fluxes:
\begin{equation}
  \Fb(\Uc) = \left ( \begin{array}{c}
         \rho u \\
         \rho X_k u \\
         \rho u^2 p \\
         (\rho E + p) u \\
         \rho e u \end{array}\right )
\end{equation}
We can write the system in conservative form for all state variables aside from $(\rho e)$ as:
\begin{equation}
  \ddt{\Uc} + \ddx{\Fb(\Uc)} = \Sc(\Uc)
\end{equation}
where for the special case of $(\rho e)$, we have an additional ``pdV'' term:
\begin{equation}
\ddt{(\rho e)} + \ddx{F(\rho e)} + p \ddx{u} = S(\rho e)
\end{equation}

@@

\begin{figure}
\centering
\plotone{strang_subch2_cfl0.8_state}
\caption{\label{fig:subch_strang_cfl08_I}}
\end{figure}

\begin{figure}
\centering
\plotone{strang_subch2_cfl0.8_mass_fractions}
\caption{\label{fig:subch_strang_cfl08_II}}
\end{figure}

@@

\begin{figure}
\centering
\plotone{sdc_subch2_cfl0.8_state}
\caption{\label{fig:subch_sdc_cfl08_I}}
\end{figure}

\begin{figure}
\centering
\plotone{sdc_subch2_cfl0.8_mass_fractions}
\caption{\label{fig:subch_sdc_cfl08_II}}
\end{figure}

@@

>>>>

\author{Melissa Rasmussen}

@@

We solve the Euler equations for compressible, reacting flow.  For ease
of exposition we describe the one-dimensional case;
multidimensional extensions are a straightforward modification to
include the CTU hydrodynamics scheme.  Our conserved variables are
\begin{equation}
  \Uc = \left ( \begin{array}{c}
           \rho \\
           \rho X_k \\
           \rho u \\
           \rho E \\
           \rho e \end{array}\right )
\end{equation}
where $\rho$ is the mass density, $u$ is velocity, $E$ is specific
total energy, $p$ is the pressure, and we carry nuclear species mass
fractions, $X_k$.  The specific total
energy relates to the specific internal energy, $e$, as $E = e + u^2/2$,
and we also separately evolve $e$, as part of a dual
energy formulation (see \citealt{bryan:1995,wdmergerI}).
 The mass fractions are constrained to sum to 1, $\sum_k X_k = 1$, but
 no such constraint exists on the auxiliary variables. 
Defining the hydrodynamical fluxes:
\begin{equation}
  \Fb(\Uc) = \left ( \begin{array}{c}
         \rho u \\
         \rho X_k u \\
         \rho u^2 p \\
         (\rho E + p) u \\
         \rho e u \end{array}\right )
\end{equation}
We can write the system in conservative form for all state variables aside from $(\rho e)$ as:
\begin{equation}
  \ddt{\Uc} + \ddx{\Fb(\Uc)} = \Sc(\Uc)
\end{equation}
where for the special case of $(\rho e)$, we have an additional ``pdV'' term:
\begin{equation}
\ddt{(\rho e)} + \ddx{F(\rho e)} + p \ddx{u} = S(\rho e)
\end{equation}

@@

\begin{figure}
\centering
\plotone{strang_subch2_cfl0_8_state}
\caption{\label{fig:subch_strang_cfl08_I}}
\end{figure}

\begin{figure}
\centering
\plotone{strang_subch2_cfl0_8_mass_fractions}
\caption{\label{fig:subch_strang_cfl08_II}}
\end{figure}

@@

\begin{figure}
\centering
\plotone{sdc_subch2_cfl0_8_state}
\caption{\label{fig:subch_sdc_cfl08_I}}
\end{figure}

\begin{figure}
\centering
\plotone{sdc_subch2_cfl0_8_mass_fractions}
\caption{\label{fig:subch_sdc_cfl08_II}}
\end{figure}

@@

\section{Reaction Network}

Our reaction network includes the nuclei: $\mathrm{n}$,
${}^{1}\mathrm{H}$, ${}^{4}\mathrm{He}$, ${}^{12}\mathrm{C}$,
${}^{14}\mathrm{C}$, ${}^{13}\mathrm{N}$, ${}^{14}\mathrm{N}$,
${}^{16}\mathrm{O}$, ${}^{18}\mathrm{O}$, ${}^{18}\mathrm{F}$,
${}^{20}\mathrm{Ne}$, ${}^{21}\mathrm{Ne}$, ${}^{22}\mathrm{Ne}$,
${}^{22}\mathrm{Na}$, ${}^{23}\mathrm{Na}$, ${}^{23}\mathrm{Mg}$,
${}^{24}\mathrm{Mg}$, ${}^{27}\mathrm{Al}$, ${}^{27}\mathrm{Si}$,
${}^{28}\mathrm{Si}$, ${}^{31}\mathrm{P}$, ${}^{31}\mathrm{S}$,
${}^{32}\mathrm{S}$, ${}^{35}\mathrm{Cl}$, ${}^{36}\mathrm{Ar}$,
${}^{39}\mathrm{K}$, ${}^{40}\mathrm{Ca}$, ${}^{43}\mathrm{Sc}$,
${}^{44}\mathrm{Ti}$, ${}^{47}\mathrm{V}$, ${}^{48}\mathrm{Cr}$,
${}^{51}\mathrm{Mn}$, ${}^{52}\mathrm{Fe}$, ${}^{55}\mathrm{Co}$,
${}^{56}\mathrm{Ni}$.

The rates linking them are:

\startlongtable
\begin{deluxetable}{ll}
\tablecaption{\label{table:react_converge_strang} Convergence ($L_1$ norm) for the reacting convergence problem using Strang splitting.}
\tablehead{\colhead{forward reaction} & \colhead{reverse reaction}}
\startdata
${}^{4}\mathrm{He} + {}^{4}\mathrm{He} + {}^{4}\mathrm{He} \rightarrow {}^{12}\mathrm{C}$ & 
  ${}^{12}\mathrm{C} \rightarrow {}^{4}\mathrm{He} + {}^{4}\mathrm{He} + {}^{4}\mathrm{He}$ \\
${}^{12}\mathrm{C} + {}^{1}\mathrm{H} \rightarrow {}^{13}\mathrm{N}$ & 
  ${}^{13}\mathrm{N} \rightarrow {}^{1}\mathrm{H} + {}^{12}\mathrm{C}$ \\
${}^{12}\mathrm{C} + {}^{4}\mathrm{He} \rightarrow {}^{16}\mathrm{O}$ & 
  ${}^{16}\mathrm{O} \rightarrow {}^{4}\mathrm{He} + {}^{12}\mathrm{C}$ \\
${}^{12}\mathrm{C} + {}^{12}\mathrm{C} \rightarrow {}^{4}\mathrm{He} + {}^{20}\mathrm{Ne}$ & 
  ${}^{20}\mathrm{Ne} + {}^{4}\mathrm{He} \rightarrow {}^{12}\mathrm{C} + {}^{12}\mathrm{C}$ \\
${}^{12}\mathrm{C} + {}^{12}\mathrm{C} \rightarrow {}^{1}\mathrm{H} + {}^{23}\mathrm{Na}$ & 
  ${}^{23}\mathrm{Na} + {}^{1}\mathrm{H} \rightarrow {}^{12}\mathrm{C} + {}^{12}\mathrm{C}$ \\
${}^{14}\mathrm{C} + {}^{4}\mathrm{He} \rightarrow {}^{18}\mathrm{O}$ & 
  ${}^{18}\mathrm{O} \rightarrow {}^{4}\mathrm{He} + {}^{14}\mathrm{C}$ \\
${}^{14}\mathrm{C} \rightarrow {}^{14}\mathrm{N}$ & 
                                         \\
${}^{13}\mathrm{N} + {}^{4}\mathrm{He} \rightarrow {}^{1}\mathrm{H} + {}^{16}\mathrm{O}$ & 
  ${}^{16}\mathrm{O} + {}^{1}\mathrm{H} \rightarrow {}^{4}\mathrm{He} + {}^{13}\mathrm{N}$ \\
${}^{14}\mathrm{N} + {}^{4}\mathrm{He} \rightarrow {}^{18}\mathrm{F}$ & 
  ${}^{18}\mathrm{F} \rightarrow {}^{4}\mathrm{He} + {}^{14}\mathrm{N}$ \\
${}^{16}\mathrm{O} + {}^{4}\mathrm{He} \rightarrow {}^{20}\mathrm{Ne}$ & 
  ${}^{20}\mathrm{Ne} \rightarrow {}^{4}\mathrm{He} + {}^{16}\mathrm{O}$ \\
${}^{16}\mathrm{O} + {}^{12}\mathrm{C} \rightarrow {}^{4}\mathrm{He} + {}^{24}\mathrm{Mg}$ & 
  ${}^{24}\mathrm{Mg} + {}^{4}\mathrm{He} \rightarrow {}^{12}\mathrm{C} + {}^{16}\mathrm{O}$ \\
${}^{16}\mathrm{O} + {}^{12}\mathrm{C} \rightarrow {}^{1}\mathrm{H} + {}^{27}\mathrm{Al}$ & 
  ${}^{27}\mathrm{Al} + {}^{1}\mathrm{H} \rightarrow {}^{12}\mathrm{C} + {}^{16}\mathrm{O}$ \\
${}^{16}\mathrm{O} + {}^{16}\mathrm{O} \rightarrow {}^{4}\mathrm{He} + {}^{28}\mathrm{Si}$ & 
  ${}^{28}\mathrm{Si} + {}^{4}\mathrm{He} \rightarrow {}^{16}\mathrm{O} + {}^{16}\mathrm{O}$ \\
${}^{16}\mathrm{O} + {}^{16}\mathrm{O} \rightarrow {}^{1}\mathrm{H} + {}^{31}\mathrm{P}$ & 
  ${}^{31}\mathrm{P} + {}^{1}\mathrm{H} \rightarrow {}^{16}\mathrm{O} + {}^{16}\mathrm{O}$ \\
${}^{16}\mathrm{O} + {}^{16}\mathrm{O} \rightarrow \mathrm{n} + {}^{31}\mathrm{S}$ & 
                                         \\
${}^{18}\mathrm{O} + {}^{4}\mathrm{He} \rightarrow {}^{22}\mathrm{Ne}$ & 
  ${}^{22}\mathrm{Ne} \rightarrow {}^{4}\mathrm{He} + {}^{18}\mathrm{O}$ \\
${}^{18}\mathrm{F} + {}^{4}\mathrm{He} \rightarrow {}^{1}\mathrm{H} + {}^{21}\mathrm{Ne}$ & 
  ${}^{21}\mathrm{Ne} + {}^{1}\mathrm{H} \rightarrow {}^{4}\mathrm{He} + {}^{18}\mathrm{F}$ \\
${}^{18}\mathrm{F} + {}^{4}\mathrm{He} \rightarrow {}^{22}\mathrm{Na}$ & 
  ${}^{22}\mathrm{Na} \rightarrow {}^{4}\mathrm{He} + {}^{18}\mathrm{F}$ \\
${}^{18}\mathrm{F} \rightarrow {}^{18}\mathrm{O}$ & 
                                         \\
${}^{20}\mathrm{Ne} + {}^{4}\mathrm{He} \rightarrow {}^{24}\mathrm{Mg}$ & 
  ${}^{24}\mathrm{Mg} \rightarrow {}^{4}\mathrm{He} + {}^{20}\mathrm{Ne}$ \\
${}^{20}\mathrm{Ne} + {}^{12}\mathrm{C} \rightarrow {}^{4}\mathrm{He} + {}^{28}\mathrm{Si}$ & 
  ${}^{28}\mathrm{Si} + {}^{4}\mathrm{He} \rightarrow {}^{12}\mathrm{C} + {}^{20}\mathrm{Ne}$ \\
${}^{20}\mathrm{Ne} + {}^{12}\mathrm{C} \rightarrow {}^{1}\mathrm{H} + {}^{31}\mathrm{P}$ & 
  ${}^{31}\mathrm{P} + {}^{1}\mathrm{H} \rightarrow {}^{12}\mathrm{C} + {}^{20}\mathrm{Ne}$ \\
${}^{21}\mathrm{Ne} + {}^{1}\mathrm{H} \rightarrow {}^{22}\mathrm{Na}$ & 
  ${}^{22}\mathrm{Na} \rightarrow {}^{1}\mathrm{H} + {}^{21}\mathrm{Ne}$ \\
${}^{22}\mathrm{Ne} + {}^{1}\mathrm{H} \rightarrow {}^{23}\mathrm{Na}$ & 
  ${}^{23}\mathrm{Na} \rightarrow {}^{1}\mathrm{H} + {}^{22}\mathrm{Ne}$ \\
${}^{22}\mathrm{Na} \rightarrow {}^{22}\mathrm{Ne}$ & 
                                         \\
${}^{23}\mathrm{Na} + {}^{1}\mathrm{H} \rightarrow {}^{4}\mathrm{He} + {}^{20}\mathrm{Ne}$ & 
  ${}^{20}\mathrm{Ne} + {}^{4}\mathrm{He} \rightarrow {}^{1}\mathrm{H} + {}^{23}\mathrm{Na}$ \\
${}^{23}\mathrm{Na} + {}^{1}\mathrm{H} \rightarrow {}^{24}\mathrm{Mg}$ & 
  ${}^{24}\mathrm{Mg} \rightarrow {}^{1}\mathrm{H} + {}^{23}\mathrm{Na}$ \\
${}^{23}\mathrm{Na} + {}^{4}\mathrm{He} \rightarrow {}^{27}\mathrm{Al}$ & 
  ${}^{27}\mathrm{Al} \rightarrow {}^{4}\mathrm{He} + {}^{23}\mathrm{Na}$ \\
${}^{23}\mathrm{Mg} + \mathrm{n} \rightarrow {}^{24}\mathrm{Mg}$ & 
                                         \\
${}^{24}\mathrm{Mg} + {}^{4}\mathrm{He} \rightarrow {}^{28}\mathrm{Si}$ & 
  ${}^{28}\mathrm{Si} \rightarrow {}^{4}\mathrm{He} + {}^{24}\mathrm{Mg}$ \\
${}^{27}\mathrm{Al} + {}^{1}\mathrm{H} \rightarrow {}^{4}\mathrm{He} + {}^{24}\mathrm{Mg}$ & 
  ${}^{24}\mathrm{Mg} + {}^{4}\mathrm{He} \rightarrow {}^{1}\mathrm{H} + {}^{27}\mathrm{Al}$ \\
${}^{27}\mathrm{Al} + {}^{1}\mathrm{H} \rightarrow {}^{28}\mathrm{Si}$ & 
  ${}^{28}\mathrm{Si} \rightarrow {}^{1}\mathrm{H} + {}^{27}\mathrm{Al}$ \\
${}^{27}\mathrm{Al} + {}^{4}\mathrm{He} \rightarrow {}^{31}\mathrm{P}$ & 
  ${}^{31}\mathrm{P} \rightarrow {}^{4}\mathrm{He} + {}^{27}\mathrm{Al}$ \\
${}^{27}\mathrm{Si} + \mathrm{n} \rightarrow {}^{28}\mathrm{Si}$ & 
                                         \\
${}^{28}\mathrm{Si} + {}^{4}\mathrm{He} \rightarrow {}^{32}\mathrm{S}$ & 
  ${}^{32}\mathrm{S} \rightarrow {}^{4}\mathrm{He} + {}^{28}\mathrm{Si}$ \\
${}^{31}\mathrm{P} + {}^{1}\mathrm{H} \rightarrow {}^{4}\mathrm{He} + {}^{28}\mathrm{Si}$ & 
  ${}^{28}\mathrm{Si} + {}^{4}\mathrm{He} \rightarrow {}^{1}\mathrm{H} + {}^{31}\mathrm{P}$ \\
${}^{31}\mathrm{P} + {}^{1}\mathrm{H} \rightarrow {}^{32}\mathrm{S}$ & 
  ${}^{32}\mathrm{S} \rightarrow {}^{1}\mathrm{H} + {}^{31}\mathrm{P}$ \\
${}^{31}\mathrm{P} + {}^{4}\mathrm{He} \rightarrow {}^{35}\mathrm{Cl}$ & 
  ${}^{35}\mathrm{Cl} \rightarrow {}^{4}\mathrm{He} + {}^{31}\mathrm{P}$ \\
${}^{31}\mathrm{S} + \mathrm{n} \rightarrow {}^{32}\mathrm{S}$ & 
                                         \\
${}^{32}\mathrm{S} + {}^{4}\mathrm{He} \rightarrow {}^{36}\mathrm{Ar}$ & 
  ${}^{36}\mathrm{Ar} \rightarrow {}^{4}\mathrm{He} + {}^{32}\mathrm{S}$ \\
${}^{35}\mathrm{Cl} + {}^{1}\mathrm{H} \rightarrow {}^{4}\mathrm{He} + {}^{32}\mathrm{S}$ & 
  ${}^{32}\mathrm{S} + {}^{4}\mathrm{He} \rightarrow {}^{1}\mathrm{H} + {}^{35}\mathrm{Cl}$ \\
${}^{35}\mathrm{Cl} + {}^{1}\mathrm{H} \rightarrow {}^{36}\mathrm{Ar}$ & 
  ${}^{36}\mathrm{Ar} \rightarrow {}^{1}\mathrm{H} + {}^{35}\mathrm{Cl}$ \\
${}^{35}\mathrm{Cl} + {}^{4}\mathrm{He} \rightarrow {}^{39}\mathrm{K}$ & 
  ${}^{39}\mathrm{K} \rightarrow {}^{4}\mathrm{He} + {}^{35}\mathrm{Cl}$ \\
${}^{36}\mathrm{Ar} + {}^{4}\mathrm{He} \rightarrow {}^{40}\mathrm{Ca}$ & 
  ${}^{40}\mathrm{Ca} \rightarrow {}^{4}\mathrm{He} + {}^{36}\mathrm{Ar}$ \\
${}^{39}\mathrm{K} + {}^{1}\mathrm{H} \rightarrow {}^{4}\mathrm{He} + {}^{36}\mathrm{Ar}$ & 
  ${}^{36}\mathrm{Ar} + {}^{4}\mathrm{He} \rightarrow {}^{1}\mathrm{H} + {}^{39}\mathrm{K}$ \\
${}^{39}\mathrm{K} + {}^{1}\mathrm{H} \rightarrow {}^{40}\mathrm{Ca}$ & 
  ${}^{40}\mathrm{Ca} \rightarrow {}^{1}\mathrm{H} + {}^{39}\mathrm{K}$ \\
${}^{39}\mathrm{K} + {}^{4}\mathrm{He} \rightarrow {}^{43}\mathrm{Sc}$ & 
  ${}^{43}\mathrm{Sc} \rightarrow {}^{4}\mathrm{He} + {}^{39}\mathrm{K}$ \\
${}^{40}\mathrm{Ca} + {}^{4}\mathrm{He} \rightarrow {}^{44}\mathrm{Ti}$ & 
  ${}^{44}\mathrm{Ti} \rightarrow {}^{4}\mathrm{He} + {}^{40}\mathrm{Ca}$ \\
${}^{43}\mathrm{Sc} + {}^{1}\mathrm{H} \rightarrow {}^{4}\mathrm{He} + {}^{40}\mathrm{Ca}$ & 
  ${}^{40}\mathrm{Ca} + {}^{4}\mathrm{He} \rightarrow {}^{1}\mathrm{H} + {}^{43}\mathrm{Sc}$ \\
${}^{43}\mathrm{Sc} + {}^{1}\mathrm{H} \rightarrow {}^{44}\mathrm{Ti}$ & 
  ${}^{44}\mathrm{Ti} \rightarrow {}^{1}\mathrm{H} + {}^{43}\mathrm{Sc}$ \\
${}^{43}\mathrm{Sc} + {}^{4}\mathrm{He} \rightarrow {}^{47}\mathrm{V}$ & 
  ${}^{47}\mathrm{V} \rightarrow {}^{4}\mathrm{He} + {}^{43}\mathrm{Sc}$ \\
${}^{44}\mathrm{Ti} + {}^{4}\mathrm{He} \rightarrow {}^{48}\mathrm{Cr}$ & 
  ${}^{48}\mathrm{Cr} \rightarrow {}^{4}\mathrm{He} + {}^{44}\mathrm{Ti}$ \\
${}^{47}\mathrm{V} + {}^{1}\mathrm{H} \rightarrow {}^{4}\mathrm{He} + {}^{44}\mathrm{Ti}$ & 
  ${}^{44}\mathrm{Ti} + {}^{4}\mathrm{He} \rightarrow {}^{1}\mathrm{H} + {}^{47}\mathrm{V}$ \\
${}^{47}\mathrm{V} + {}^{1}\mathrm{H} \rightarrow {}^{48}\mathrm{Cr}$ & 
  ${}^{48}\mathrm{Cr} \rightarrow {}^{1}\mathrm{H} + {}^{47}\mathrm{V}$ \\
${}^{47}\mathrm{V} + {}^{4}\mathrm{He} \rightarrow {}^{51}\mathrm{Mn}$ & 
  ${}^{51}\mathrm{Mn} \rightarrow {}^{4}\mathrm{He} + {}^{47}\mathrm{V}$ \\
${}^{48}\mathrm{Cr} + {}^{4}\mathrm{He} \rightarrow {}^{1}\mathrm{H} + {}^{51}\mathrm{Mn}$ & 
  ${}^{51}\mathrm{Mn} + {}^{1}\mathrm{H} \rightarrow {}^{4}\mathrm{He} + {}^{48}\mathrm{Cr}$ \\
${}^{48}\mathrm{Cr} + {}^{4}\mathrm{He} \rightarrow {}^{52}\mathrm{Fe}$ & 
  ${}^{52}\mathrm{Fe} \rightarrow {}^{4}\mathrm{He} + {}^{48}\mathrm{Cr}$ \\
${}^{51}\mathrm{Mn} + {}^{1}\mathrm{H} \rightarrow {}^{52}\mathrm{Fe}$ & 
  ${}^{52}\mathrm{Fe} \rightarrow {}^{1}\mathrm{H} + {}^{51}\mathrm{Mn}$ \\
${}^{51}\mathrm{Mn} + {}^{4}\mathrm{He} \rightarrow {}^{55}\mathrm{Co}$ & 
  ${}^{55}\mathrm{Co} \rightarrow {}^{4}\mathrm{He} + {}^{51}\mathrm{Mn}$ \\
${}^{52}\mathrm{Fe} + {}^{4}\mathrm{He} \rightarrow {}^{1}\mathrm{H} + {}^{55}\mathrm{Co}$ & 
  ${}^{55}\mathrm{Co} + {}^{1}\mathrm{H} \rightarrow {}^{4}\mathrm{He} + {}^{52}\mathrm{Fe}$ \\
${}^{52}\mathrm{Fe} + {}^{4}\mathrm{He} \rightarrow {}^{56}\mathrm{Ni}$ & 
  ${}^{56}\mathrm{Ni} \rightarrow {}^{4}\mathrm{He} + {}^{52}\mathrm{Fe}$ \\
${}^{55}\mathrm{Co} + {}^{1}\mathrm{H} \rightarrow {}^{56}\mathrm{Ni}$ & 
  ${}^{56}\mathrm{Ni} \rightarrow {}^{1}\mathrm{H} + {}^{55}\mathrm{Co}$ \\
 &  ${}^{12}\mathrm{C} + {}^{12}\mathrm{C} \rightarrow \mathrm{n} + {}^{23}\mathrm{Mg}$ \\
 &  ${}^{16}\mathrm{O} + {}^{12}\mathrm{C} \rightarrow \mathrm{n} + {}^{27}\mathrm{Si}$ \\
\enddata
\end{deluxetable}
